A fast unified model for parsing and sentence understanding

### 2016-03-15 writing/hard_stack_paper/paper.tex
Resolve todos in efficiency section

A naive implementation of the SPINN model would require simulating a stack of size $N$ for each input sentence. In order to update the model weights via backpropagation, we would also need to maintain all intermediate stack representations for later chain rule calculations. This implies a total space requirement of $M \times N \times T \times D$, which is prohibitively large for significant batch sizes $M$ or sentence lengths $N$. Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift~and \reduce~operations only affect the top of the stack). \todo{[JG] Remind reader what T and D are here.}

We designed an alternative minimal-space stack algorithm inspired by the zipper technique \citep{huet1997zipper}. For a single input sentence, we represent the stack with a single $T \times D$ matrix $S$. \todo{[JG] Either expand or remove references to batch size.} Each row $S_t$ represents the top of the actual stack at timestep $t$. We maintain a queue of backpointers onto $S$ in order to track which elements should be involved in a \reduce~operation at any given time. Algorithm~\ref{alg:thin-stack} below describes the full mechanics of a stack feedforward in this compressed representation. It operates on the compressed $T \times D$ matrix $S$ and a backpointer queue $Q$. Table~\ref{tbl:thin-stack} shows an example run of this algorithm.

>>>>

A naive implementation of the SPINN model would require simulating a stack of size $N$ for each input sentence. In order to update the model weights via backpropagation, we would also need to maintain all intermediate stack representations for later chain rule calculations. This implies a per-example space requirement of $N \times T \times D$, which is prohibitively large for significant batch sizes $M$ or sentence lengths $N$. (Recall that $T = 2N - 1$ is the number of transitions in the shift-reduce parsing sequence, and $D$ is the dimension of the vectors on the stack.) Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift~and \reduce~operations only affect the top of the stack).

We designed an alternative minimal-space stack algorithm inspired by the zipper technique \citep{huet1997zipper}. For a single input sentence, we represent the stack with a single $T \times D$ matrix $S$. Each row $S_t$ represents the top of the actual stack at timestep $t$. We maintain a queue of backpointers onto $S$ in order to track which elements should be involved in a \reduce~operation at any given time. Algorithm~\ref{alg:thin-stack} below describes the full mechanics of a stack feedforward in this compressed representation. It operates on the compressed $T \times D$ matrix $S$ and a backpointer queue $Q$. Table~\ref{tbl:thin-stack} shows an example run of this algorithm.
